Describing vocalizations in young children: A big data approach through citizen science annotation

### 2021-02-18 202010_jslhr/paper.tex
ready for final proofing

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Children's vocalizations and citizen science}
\author{Chiara Semenzin\textsuperscript{1}, Lisa Hamrick\textsuperscript{2}, Amanda Seidl\textsuperscript{2}, Bridgette Kelleher\textsuperscript{2}, \& Alejandrina Cristia\textsuperscript{1}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University\\\textsuperscript{2} Purdue University}
\authornote{AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017); and the J. S. McDonnell Foundation Understanding Human Cognition Scholar Award. LH acknowledges the National Institute of Deafness and Other Communication Disorders (F31DC018219). BK acknowledges the National Institute of Mental Health (K23MH111955) and the Kinley Trust. This publication uses data generated via the Zooniverse.org platform, development of which is funded by generous support, including a Global Impact Award from Google, and by a grant from the Alfred P. Sloan Foundation. The funders had no impact on this study.  All authors approved the final manuscript as submitted and agree to be accountable for all aspects of the work and have no conflict of interests to disclose.

Correspondence concerning this article should be addressed to Alejandrina Cristia, 29 rue d'Ulm, 75005, Paris, France. E-mail: alecristia@gmail.com}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage[titles]{tocloft}
\cftpagenumbersoff{figure}
\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
\renewcommand{\cftfigaftersnum}{.\space}
\setlength{\cftfigindent}{0pt}
\setlength{\cftafterloftitleskip}{0pt}
\settowidth{\cftfignumwidth}{Figure 10.\qquad}
\usepackage{setspace}
\AtBeginEnvironment{tabular}{\singlespacing}
\AtBeginEnvironment{lltable}{\singlespacing}
\AtBeginEnvironment{tablenotes}{\singlespacing}
\captionsetup[table]{font={stretch=1.5}}
\captionsetup[figure]{font={stretch=1.5}}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

@@

\abstract{
\textbf{Purpose:} Recording young children's vocalizations through wearables is a promising method. However, accurately and rapidly annotating these files remains challenging. Online crowdsourcing with the collaboration of citizen scientists could be a feasible solution. In this paper, we assess the extent to which citizen scientists' annotations align with those gathered in the lab for recordings collected from young children. \newline

\textbf{Method:} Segments identified by LENA\textsuperscript{TM} as produced by the key child were extracted from one daylong recording for each of 20 participants: 10 low-risk control children and 10 children diagnosed with Angelman syndrome, a neurogenetic syndrome characterized by severe language impairments. Speech samples were annotated by trained annotators in the laboratory as well as by citizen scientists on Zooniverse. All annotators assigned one of five labels to each sample: Canonical, Non-Canonical, Crying, Laughing, and Junk. This allowed the derivation of two child-level vocalization metrics: the Linguistic Proportion, and the Canonical Proportion.
\newline

@@

\textbf{Conclusion:} Annotations obtained through a citizen science platform can help us overcome challenges posed by the process of annotating daylong speech recordings. Particularly when used in composites or derived metrics, such annotations can be used to investigate early markers of language delays in non-typically developing children.
}

@@

Since early language delays can adversely affect children's literacy, behavior, social interaction, and scholastic achievement extending well into adulthood, early interventions have been described as a better societal investment than later ones (Heckman, 2006). Some research suggests that spontaneous behavior, captured for instance via home videos, could provide important indices to development (Belardi et al., 2017; Overby, Belardi, \& Schreiber, 2020). Advancements in the field of wearable technologies, such as LENA\textsuperscript{TM} recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally (Oller et al., 2010; Rankine et al., 2017; VanDam \& Yoshinaga-Itano, 2019). Wearable recorders allow data collection to happen in the child's natural environment, and at a large scale, which may be particularly helpful for children whose speech is not easily elicited. Although such long-form recordings are increasingly common (Ganek \& Eriks-Brophy, 2018), challenges remain with respect to how these data are handled, annotated, and analyzed (Casillas \& Cristia, 2019). The present work reports on the validity of labels and child-level descriptors of children's vocalizations gathered from citizen scientists, in a comparison to expert lab annotators.

@@

A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis, and it is beyond the scope of this paper to provide a full summary of this work {[}but see \textbf{XX ADD REFERENCE} for a recent summary{]}. Moreover, the way in which vocal development has been studied in the past may not easily translate to long-form recordings. As mentioned earlier, long-form recordings have several advantages, including capturing the child's vocal patterns in their natural environment and being able to accumulate a great amount of data easily, which is particularly useful for diagnoses and pathologies characterized by low levels of vocal production. However, this also means that the audiorecording is harder to process than an audiorecording gathered in more manicured and stable conditions: In a typical daylong recording, the child will sometimes be in a quiet room with just her primary caregiver, but later in a noisy supermarket or having dinner with family and friends. In fact, human annotation of such audio is estimated to take about 30 times the audio length, with these estimates being greater the more precise one wants or needs to be (Casillas et al., 2017). It is for this reason that most users of long-form recordings have turned to automatized software for at least a first-pass analysis.

@@

The most commonly used software was created by the LENA\textsuperscript{TM} Foundation, and it returns a diarization of the audio signal split into key talkers, including the child wearing the device, as well as a split of these audio attributed to the child into three categories: crying, other fixed and vegetative signals, and speech vocalizations. These subcategories have not been widely validated (Cristia et al., 2020). Moreover, LENA\textsuperscript{TM} lacks an important distinction between more and less advanced vocalizations.

Other automatized algorithms have been developed in the last two years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical (Schuller et al., 2019). The split of speech vocalizations into canonical and non-canonical, and the overt inclusion of laughing, allow a somewhat finer-grained picture of children's vocal activity than what the LENA\textsuperscript{TM} algorithm currently offers. Another advantage of this new line of research is that there exists a challenge where performance has been benchmarked, so unlike LENA\textsuperscript{TM}'s algorithms, we know which algorithms perform better than others. Indeed, the ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with a test-set unweighted average recall of around 55\%. By re-using the same test set, researchers can develop new algorithms and prove that they can outperform this baseline (or any subsequent best-performing algorithm benchmarked in the same test set).

Table 1 provides an idea of the accuracy across different methods. Specifically, we calculated accuracy of the LENA\textsuperscript{TM}'s subtypes\footnote{If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as \enquote{Junk} or not categorized.} based on the laboratory annotations used in the present study; and we also show the results of ComParE 2019 BabySounds sub-challenge, although note it is not based on the same data as the other two columns. Table 1 shows that LENA\textsuperscript{TM} has a good recall for speech-like vocalizations and crying but not laughing, and by design it does not distinguish between canonical and non-canonical vocalizations. The ComParE 2019 BabySounds sub-challenge's baseline model achieved a more moderate recall on crying than LENA, but clearly outperforms the LENA\textsuperscript{TM} algorithm in the Laughing category, with the additional advantage of distinguishing canonical and non-canonical -- although note that performance in non-canonical is barely above a chance level of 20\% (based on 5 categories, assuming equal probabilities). Notice that the baseline model also included a category called \enquote{Junk} for sections that turned out not to be the child's vocalizations after all, but LENA\textsuperscript{TM} here has a low recall because only sections that LENA\textsuperscript{TM} classified as being child vocalizations are considered in the present study, and thus this class's category cannot but be low-performing for LENA\textsuperscript{TM}. That is, since we selected only segments attributed to the key child, any segment that turns out not to be a child vocalization is a LENA\textsuperscript{TM} error. We return to the comparison between LENA\textsuperscript{TM}, the ComParE 2019 BabySounds sub-challenge baseline, and our Zooniverse-based method in the Discussion, but for now, we simply point out that there are new algorithms being developed, but their performance --although competitive when compared to LENA\textsuperscript{TM}-- is underwhelming.

@@

One possibility that has only recently begun to be explored is the use of derived metrics. So instead of evaluating algorithms and other annotation procedures on their accuracy at individual labels, one can derive a metric that more closely relates to vocalization development.

For example, most work using LENA\textsuperscript{TM} software reports on children's vocalization counts -- this is a derived metric because it is not simply a description of sections of the audio, but instead it integrates over a certain period of time. Although this has been sometimes criticized as being more about quantity than quality (McDaniel, Yoder, Estes, \& Rogers, 2020), this is a promising metric of individuals' vocalization development because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA\textsuperscript{TM} (Cristia et al., 2020). Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r \textasciitilde{} .3 with standardized language scores in a meta-analysis (Wang, Williams, Dilley, \& Houston, 2020). Thus despite representing a composite of skills, vocalization count may provide a useful estimate of vocalization development.

Here, we focus on two alternative composite metrics. One metric that can be derived once children's vocalizations are split into crying, laughing, canonical, and non-canonical is the Linguistic Proportion: the proportion of vocalizations that are linguistic (canonical and non-canonical) out of all vocalizations. To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age.

Yet another metric that could be extracted to estimate relative linguistic complexity is the Canonical Proportion: the proportion of vocalizations that contain a canonical transition or syllable out of all linguistic (canonical and non-canonical) vocalizations. One recent study has established that the Canonical Proportion extracted from daylong recordings is significantly correlated with age in a multicultural and multilingual sample (Cychosz et al., 2019). Importantly, this association was established in a sample of children going up to 3 years of age, and thus well beyond the babbling period, and into first words and word combinations.

There is a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this work focuses on babbling and thus typically on infants under one year of age. A critical milestone involves the increasingly common production of canonical syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech (Oller, Eilers, Neal, \& Cobo-Lewis, 1998). Given its adult-like consonant-vowel or vowel-consonant structure, canonical babble is considered to be a starting point on the path to recognizable speech. Canonical syllables show a higher complexity given the smooth articulatory transition between a consonant and a vowel (or vice versa), when compared to more primitive sounds such as squeals, or isolated vowels. Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development (Oller, 2000). In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization counts in a sample of children diagnosed with an autism spectrum disorder (McDaniel et al., 2020).

@@

In sum, there is a growing literature attempting to use data from daylong recordings to describe young children's vocalizations, but there are two outstanding challenges. The first pertains to how the data are annotated, with human annotation being costly. The second relates to how descriptors of vocal development are generated, i.e., how annotations of audio sections are integrated into vocal development metrics.

Crowdsourcing refers to the process whereby a task is solved by a crowd, rather than an individual. A number of fields, particularly in the data-driven sciences, have already engaged in the collection of data (including annotations) through crowdsourcing, thanks to its low cost and ecological value (Crump, McDonnell, \& Gureckis, 2013; Sescleifer, Francoisse, \& Lin, 2018). For example, a systematic review on crowd-sourced ratings of speech found that \enquote{lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology}; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81 (Sescleifer et al., 2018). On the other hand, the systematic review returned only 8 studies (of which only four were published in peer-reviewed journals), suggesting that there is considerable need for further research on this topic.

Mechanical Turk (MTurk) MTurk is an online labor market created by Amazon to assist \enquote{requesters} in hiring and paying \enquote{workers} for the completion of computerized tasks. Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable (Berinsky, Huber, \& Lenz, 2012), some question marks are raised as some \enquote{workers} turn out to be bots, or are poorly motivated (and potentially exploited) humans.

@@

Citizen science may be particularly helpful when analyzing infants' data from wearables, which remain challenging to annotate as mentioned above. There is one previous study that attempted this approach. Cychosz et al. (2019) drew child vocalization data from a diverse set of corpora centered on children learning one of four languages: English, Tseltal, Tsimane', and Yélî. For the English and Tsimane' corpora, vocalizations were automatically identified using LENA\textsuperscript{TM}, whereas the other two were extracted through manual segmentation. Segments were then split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received three classifications into Crying, Laughing, Canonical, Non-Canonical, or Junk (with the latter tag used for clips that did not contain a child voice). They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgment of canonical or non-canonical. They found that this Canonical Proportion increased with infant age in an analysis collapsing across corpora, which provides a first proof of principle that citizen scientists' annotations could be an appropriate solution to the problem of annotating child vocalizations from wearables. The authors also studied how their Canonical Proportion related to age within each corpus that had a sufficient number of children varying in age. They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the \enquote{Junk} category was very high. Additionally, they found that the correlation between Canonical Proportion and age was much weaker in two corpora where vocalizations were segmented with LENA\textsuperscript{TM} than through manual segmentation. In sum, these analyses within corpora reveal that data may not be uniformly useful, with potential variation across corpora, children, and methods.

@@

Our work seeks to broaden the already promising results that have emerged from previous work summarized above. We hope that this methodology will open the road to larger scale analyses of children's vocalizations as captured by wearables. In addition, we went beyond Cychosz et al. (2019)\enquote{s study in two important ways. First, we relied on the largest and best established citizen science platform: Zooniverse hosts more than 1.6 million users from diverse walks of life, and it offers a completely automatized API system to more easily scale tasks in a transparent and cumulative science fashion. In contrast, Cychosz et al. (2019) actively recruited people to provide annotations, stating that \enquote{{[}a{]}nnotators included language and speech researchers, undergraduate students, research assistants, and other interested parties, totaling 136 unique annotators}. A process that draws from research assistants relies on expert resources, whose time is expensive. Moreover, the people these researchers recruit tend to be more accustomed to child vocalization data, which may improve their annotation performance, and thus provide an overestimate of the quality of the annotations that can be realistically done within a truly citizen science framework. Those authors also had to rely on manual extraction of the resulting data from the iHearUplay team, which added to the administrative load. Second, a core goal of the present work was to determine how citizen scientists} annotations fare compared to the current gold standard, laboratory annotations. Cychosz et al. (2019) did not have a gold standard for their data.

@@

When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk judgments); and the Canonical Proportion (Canonical)/(all linguistic judgments).

It was important for generalizability purposes to examine data from children with variable biological, mental age, and vocal maturity levels. We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA\textsuperscript{TM} (see meta-analyses in Wang et al., 2020; Cristia et al., 2020); as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range. Some previous work had shown that vocal development proceeded differently when compared to a normative sample (Hamrick et al., 2019; Hamrick \& Tonnsen, 2019). Generally, studies on vocal development among children with genetic disorders reveal lower volubility and lower prevalence of canonical syllables (Belardi et al., 2017; Rankine et al., 2017; see also Roche et al., 2018 for a review of retrospective video analysis of children diagnosed with Autism Spectrum Disorder, Rett Syndrome, or Fragile X, documenting these and other vocal characteristics). Low volubility is one of the contexts in which daylong recordings may be particularly advantageous over alternative recording methods (see Rankine et al., 2017 for discussion), since it is challenging to measure vocal development accurately based on shorter recordings (as the child may not vocalize at all in a 1-2h recording). Additionally, the ideal method will be sensitive not only to the large changes that occur over normative development, but also to the potentially smaller individual differences found among children. Such a level of sensitivity is important if these methods are ever to be used to detect potential effects of intervention.

@@

All analyses and key data are available for reproduction on OSF (\url{https://osf.io/57yha/}). Computer code used to collect the Zooniverse judgments is available from GitHub (Semenzin \& Cristia, 2020). This manuscript is reproducible thanks to the use of RMd (Baumer \& Udwin, 2015) and Papaja (Aust \& Barth, 2017) on R (Team \& others, 2013).

>>>>

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Children's vocalizations and citizen science}
\author{Chiara Semenzin\textsuperscript{1}, Lisa Hamrick\textsuperscript{2}, Amanda Seidl\textsuperscript{2}, Bridgette Kelleher\textsuperscript{2}, \& Alejandrina Cristia\textsuperscript{1}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University\\\textsuperscript{2} Purdue University}
\authornote{AC acknowledges Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004 ACLEW, ANR-14-CE30-0003 MechELex, ANR-17-EURE-0017); and the J. S. McDonnell Foundation (Understanding Human Cognition Scholar Award). LH acknowledges the National Institute of Deafness and Other Communication Disorders (F31DC018219). BK acknowledges the National Institute of Mental Health (K23MH111955) and the Kinley Trust. This publication uses data generated via the Zooniverse.org platform, development of which is funded by generous support, including a Global Impact Award from Google, and by a grant from the Alfred P. Sloan Foundation. The funders had no impact on this study.  All authors approved the final manuscript as submitted, agree to be accountable for all aspects of the work and have no conflict of interests to disclose.

Correspondence concerning this article should be addressed to Alejandrina Cristia, 29 rue d'Ulm, 75005, Paris, France. E-mail: alecristia@gmail.com}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\usepackage[titles]{tocloft}
\cftpagenumbersoff{figure}
\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
\renewcommand{\cftfigaftersnum}{.\space}
\setlength{\cftfigindent}{0pt}
\setlength{\cftafterloftitleskip}{0pt}
\settowidth{\cftfignumwidth}{Figure 10.\qquad}
\usepackage{setspace}
\AtBeginEnvironment{tabular}{\singlespacing}
\AtBeginEnvironment{lltable}{\singlespacing}
\AtBeginEnvironment{tablenotes}{\singlespacing}
\captionsetup[table]{font={stretch=1.5}}
\captionsetup[figure]{font={stretch=1.5}}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

@@

\abstract{
\textbf{Purpose:} Recording young children's vocalizations through wearables is a promising method to assess language development. However, accurately and rapidly annotating these files remains challenging. Online crowdsourcing with the collaboration of citizen scientists could be a feasible solution. In this paper, we assess the extent to which citizen scientists' annotations align with those gathered in the lab for recordings collected from young children. \newline

\textbf{Method:} Segments identified by LENA® as produced by the key child were extracted from one daylong recording for each of 20 participants: 10 low-risk control children and 10 children diagnosed with Angelman syndrome, a neurogenetic syndrome characterized by severe language impairments. Speech samples were annotated by trained annotators in the laboratory as well as by citizen scientists on Zooniverse. All annotators assigned one of five labels to each sample: Canonical, Non-Canonical, Crying, Laughing, and Junk. This allowed the derivation of two child-level vocalization metrics: the Linguistic Proportion, and the Canonical Proportion.
\newline

@@

\textbf{Conclusion:} Annotations obtained through a citizen science platform can help us overcome challenges posed by the process of annotating daylong speech recordings. Particularly when used in composites or derived metrics, such annotations can be used to investigate early markers of language delays.
}

@@

Since early language delays can adversely affect children's literacy, behavior, social interaction, and scholastic achievement extending well into adulthood, early interventions have been described as a better societal investment than later ones (Heckman, 2006). Some research suggests that spontaneous behavior, captured for instance via home videos, could provide important indices to development (Belardi et al., 2017; Overby, Belardi, \& Schreiber, 2020). Advancements in the field of wearable technologies, such as LENA® recorders, have opened new avenues to both early detection of speech pathologies and research in language development more generally (Oller et al., 2010; Rankine et al., 2017; VanDam \& Yoshinaga-Itano, 2019). Wearable recorders allow data collection to happen in the child's natural environment, and at a large scale, which may be particularly helpful for children whose speech is not easily elicited. Although such long-form recordings are increasingly common (Ganek \& Eriks-Brophy, 2018), challenges remain with respect to how these data are handled, annotated, and analyzed (Casillas \& Cristia, 2019). The present work reports on the validity of labels and child-level descriptors of children's vocalizations gathered from citizen scientists, in a comparison to expert lab annotators.

@@

A large body of research has investigated both fine-grained and coarse descriptions of children's vocalizations as a function of age and diagnosis. It is beyond the scope of this paper to provide a full summary of this extensive body work, mainly because the way in which vocal development has been studied in the past may not easily translate to long-form recordings. Long-form recordings have several advantages, including capturing the child's vocal patterns in their natural environment and being able to accumulate a great amount of data easily, which is particularly useful for diagnoses and pathologies characterized by low levels of vocal production. However, this also means that the audiorecording is harder to process than an audiorecording gathered in more manicured and stable conditions: In a typical daylong recording, the child will sometimes be in a quiet room with just her primary caregiver, but later in a noisy supermarket or having dinner with family and friends. In fact, human annotation of such audio is estimated to take about 30 times the audio length, with these estimates being greater the more precise one wants or needs to be (Casillas et al., 2017). It is for this reason that most users of long-form recordings have turned to automatized software for at least a first-pass analysis.

@@

The most commonly used software was created by the LENA® Foundation, and it returns a diarization of the audio signal split into key talkers, including the child wearing the device, as well as a split of sections attributed to the child into three categories: crying, other fixed and vegetative signals, and speech vocalizations. These subcategories have not been widely validated (Cristia et al., 2020). Moreover, until recently LENA® lacked an important distinction between more and less advanced vocalizations.\footnote{The latest version of the LENA® software does tag canonical utterances separately from non-canonical ones.}

Other automatized algorithms have been developed in the last two years, saliently ones attempting to classify child vocalizations into crying, laughing, canonical, and non-canonical (Schuller et al., 2019). The overt inclusion of laughing allows a somewhat finer-grained picture of children's vocal activity than what the LENA® algorithm currently offers. Another advantage of this new line of research is that there exists a challenge where performance has been benchmarked, so unlike LENA®'s algorithms, we know which algorithms perform better than others. Indeed, the ComParE 2019 BabySounds sub-challenge established a state-of-the-art baseline with an unweighted average recall of around 55\% on the test set. By re-using the same test set, researchers can develop new algorithms and prove that they can outperform this baseline (or any subsequent best-performing algorithm benchmarked in the same test set).

Table 1 provides an idea of the accuracy across different methods. Specifically, we calculated accuracy of the LENA®'s subtypes\footnote{If a child segment contained speech, then it counted towards canonical/non-canonical; else, if it contained crying it counted towards crying; else, if it contained some fixed signals it counted towards laughing; a small proportion were left that did not have any of the three and were considered as \enquote{Junk} or not categorized. When these data were analyzed, the LENA® software did not distinguish canonical and non-canonical.} based on the laboratory annotations used in the present study; and we also show the results of ComParE 2019 BabySounds sub-challenge, although note it is not based on the same data as the other two columns. Table 1 shows that LENA® has a good recall for speech-like vocalizations and crying but not laughing, and at the time it did not distinguish between canonical and non-canonical vocalizations. The ComParE 2019 BabySounds sub-challenge's baseline model achieved a more moderate recall on crying than LENA, but clearly outperforms the LENA® algorithm in the Laughing category, with the additional advantage of distinguishing canonical and non-canonical -- although note that performance in non-canonical is barely above a chance level of 20\% (based on 5 categories, assuming equal probabilities). Notice that the baseline model also included a category called \enquote{Junk} for sections that turned out not to be the child's vocalizations after all. LENA® has a low recall for the Junk category because only sections that LENA® classified as being child vocalizations are considered in the present analysis. As a result, any segment that turns out not to be a child vocalization is a LENA® error. We return to the comparison between LENA®, the ComParE 2019 BabySounds sub-challenge baseline, and our Zooniverse-based method in the Discussion; for now, we simply point out that there are new algorithms being developed, but their performance --although competitive when compared to LENA®-- is underwhelming.

@@

One possibility that has only recently begun to be explored is the use of \emph{derived metrics}. So instead of evaluating algorithms and other annotation procedures on accuracy of individual sections of the audio, one can integrate over time (e.g., over the whole day-long audio) to derive a metric that more closely relates to vocalization development.

For example, most work using LENA® software reports on children's vocalization counts -- a derived metric because it is not simply a description of sections of the audio, but instead it integrates over the whole recording length. Although linguistic vocalization count has been criticized as being more about quantity than quality (McDaniel, Yoder, Estes, \& Rogers, 2020), it is a promising metric of individuals' vocalization development because it shows correlations with age (which is a proxy of development), and it can be extracted quite accurately with LENA® (Cristia et al., 2020). Additionally, the child vocalization count metric has been found to be concurrently and predictively correlated with an effect size r \textasciitilde{} .3 with standardized language scores in a meta-analysis (Wang, Williams, Dilley, \& Houston, 2020). LENA®'s child vocalization counts (CVC) aggregates all key child vocalizations detected as being \emph{linguistic}. As a result, an increase in CVC could indicate an actual increase in how many linguistic vocalizations the child produces, or a change in how well they are detected as such by the algorithm. Thus, despite representing a composite of skills, vocalization count may provide a useful estimate of vocalization development.

Here, we focus on two alternative composite metrics. One metric that can be derived once children's vocalizations are split into crying, laughing, canonical, and non-canonical is the Linguistic Proportion: the proportion of vocalizations that are linguistic (canonical and non-canonical) out of all vocalizations. To our knowledge, this metric has not been extensively explored previously, but it is likely that this proportion increases with age early on and eventually plateaus.

Yet another metric that could be extracted to estimate relative linguistic complexity is the Canonical Proportion: the proportion of vocalizations that contain a canonical transition or syllable out of all linguistic (canonical and non-canonical) vocalizations. Cychosz et al. (2021) found that the Canonical Proportion extracted from daylong recordings was significantly correlated with age in a multicultural and multilingual sample of children up to 3 years of age, and thus well beyond the babbling period, and into first words and word combinations. That study built on a wider research base documenting the potential importance of the Canonical Proportion, and related metrics, although this prior work focuses on babbling and thus typically on infants under one year of age. In these younger infants, a critical milestone involves the increasingly common production of canonical syllables, consonant-vowel or vowel-consonant sequences that resemble those found in adult speech (Oller, Eilers, Neal, \& Cobo-Lewis, 1998). Given its adult-like consonant-vowel or vowel-consonant structure, canonical babble is considered to be a starting point on the path to recognizable speech. When compared to more primitive sounds such as squeals, or isolated vowels, canonical syllables show a higher complexity given the smooth articulatory transition between a consonant and a vowel (or vice versa). Some work suggests that Canonical Proportions above .15 are expected by about 10 months of age in typical development (Cychosz et al., 2021; Oller, 2000). In addition, the proportion of vocalizations containing a canonical syllable has been found to be more predictive of individual development than sheer vocalization counts in a sample of children diagnosed with an autism spectrum disorder (McDaniel et al., 2020).

@@

In sum, there is a growing literature attempting to use data from daylong recordings to describe young children's vocalizations, but there are two outstanding challenges. The first pertains to how the data are annotated, with human annotation being costly. The second relates to how descriptors of vocal development are generated, i.e., how annotations of individual audio sections are integrated over all data for a given child to derive child-level vocal development metrics.

\enquote{Crowdsourcing} refers to the process whereby a task is solved by a crowd, rather than an individual. A number of fields, particularly in the data-driven sciences, have already engaged in the collection of data (including annotations) through crowdsourcing, thanks to its low cost and ecological value (Crump, McDonnell, \& Gureckis, 2013; Sescleifer, Francoisse, \& Lin, 2018). For example, a systematic review on crowdsourced ratings of speech found that \enquote{lay ratings are highly concordant with expert opinion, validating crowdsourcing as a reliable methodology}; across studies, crowdsourced and expert listener classifications yielded a mean correlation coefficient of .81 (Sescleifer et al., 2018). On the other hand, the systematic review returned only 8 studies (of which only four were published in peer-reviewed journals), suggesting that there is considerable need for further research on this topic.

Mechanical Turk (MTurk) MTurk is an online labor market created by Amazon to assist \enquote{requesters} in hiring and paying \enquote{workers} for the completion of computerized tasks. Although it is a leading crowdsourcing service, and some evidence suggests MTurkers' annotations can be quite reliable (Berinsky, Huber, \& Lenz, 2012), question marks are raised as some \enquote{workers} turn out to be bots, or are poorly motivated (and potentially exploited) humans.

@@

Citizen science may be particularly helpful when analyzing infants' data from wearables, which remain challenging to annotate as mentioned above. There is one previous study that attempted this approach. Cychosz et al. (2021) drew child vocalization data from a diverse set of corpora centered on children learning one of four languages: English, Tseltal, Tsimane', and Yélî. For the English and Tsimane' corpora, vocalizations were automatically identified using LENA®, whereas the other two were extracted through manual segmentation. Segments were then split into maximally 500-ms long clips, and presented to annotators through the citizen science iHearUPlay platform, with the aim that each clip received three classifications into Crying, Laughing, Canonical, Non-Canonical, or Junk (with the latter tag used for clips that did not contain a child voice). They then derived an implementation of the Canonical Proportion, as the proportion of clips receiving a majority judgment of canonical out of the clips receiving a majority judgment of canonical or non-canonical. They found that this Canonical Proportion increased with infant age in an analysis collapsing across corpora, which provides a first proof of principle that citizen scientists' annotations could be an appropriate solution to the problem of annotating child vocalizations from wearables. The authors also studied how their Canonical Proportion related to age within each corpus that had a sufficient number of children varying in age. They found that in one English corpus bearing on very young infants, the proportion of clips assigned to the \enquote{Junk} category was very high. Additionally, they found that the correlation between Canonical Proportion and age was much weaker in two corpora where vocalizations were segmented with LENA® than through manual segmentation. In sum, these analyses within corpora reveal that data may not be uniformly useful, with potential variation across corpora, children, and methods.

@@

Our work seeks to broaden the already promising results that have emerged from previous work summarized above. We hope that this methodology will open the road to larger scale analyses of children's vocalizations as captured by wearables. In addition, we went beyond Cychosz et al. (2021)\enquote{s study in two important ways. First, we relied on the largest and best established citizen science platform: Zooniverse hosts more than 1.6 million users from diverse walks of life, and it offers a completely automatized API system to more easily scale tasks in a transparent and cumulative science fashion. In contrast, Cychosz et al. (2021) actively recruited people to provide annotations, stating that \enquote{{[}a{]}nnotators included language and speech researchers, undergraduate students, research assistants, and other interested parties, totaling 136 unique annotators}. A process that draws from research assistants relies on expert resources, whose time is expensive. Moreover, the people these researchers recruit tend to be more accustomed to child vocalization data, which may improve their annotation performance, and thus provide an overestimate of the quality of the annotations that can be realistically done within a truly citizen science framework. Second, a core goal of the present work was to determine how citizen scientists} annotations fare compared to the current gold standard, laboratory annotations. Cychosz et al. (2021) did not have a gold standard for their data.

@@

When considering correspondence at the level of individual children, we derived two metrics: the Linguistic Proportion (Canonical+Non-Canonical)/(all non-Junk labels); and the Canonical Proportion (Canonical)/(all linguistic labels).

It was important for generalizability purposes to examine data from children with variable biological age, mental age, and vocal maturity levels. We therefore included a group of low-risk control infants, who are close to the population most commonly sampled in studies using LENA® (see meta-analyses in Wang et al., 2020; Cristia et al., 2020); as well as children who had been diagnosed with Angelman syndrome and whose ages spanned a wide age range. Some previous work had shown that vocal development proceeded differently among children with an Angelman syndrome diagnosis when compared to a normative sample (Hamrick et al., 2019; Hamrick \& Tonnsen, 2019). Generally, studies on vocal development among children with genetic disorders reveal lower volubility and lower prevalence of canonical syllables (Belardi et al., 2017; Rankine et al., 2017; see also Roche et al., 2018 for a review of retrospective video analysis of children diagnosed with Autism Spectrum Disorder, Rett Syndrome, or Fragile X, documenting these and other vocal characteristics). Low volubility is one of the contexts in which daylong recordings may be particularly advantageous over alternative recording methods (see Rankine et al., 2017 for discussion), since it is challenging to measure vocal development accurately based on shorter recordings (as the child may not vocalize at all in a 1-2h recording). Additionally, the ideal method is sensitive not only to the large changes that occur over normative development, but also to the potentially smaller individual differences found among children. Such a level of sensitivity is important if these methods are ever to be used to detect potential effects of intervention.

@@

All analyses and key data are available for reproduction on OSF (\url{https://osf.io/57yha/}). Computer code used to collect the Zooniverse judgments is available from GitHub (Semenzin \& Cristia, 2020). This manuscript is reproducible thanks to the use of RMd (Baumer \& Udwin, 2015) and Papaja (Aust \& Barth, 2017) on R (R Core Team \& others, 2013).
